State Results

### 2023-06-30 reports/state_figures.Rmd
figures

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)

@@

@@

# read state-level results
targets_dir <- here("_targets_6_26")

@@

state_res <- state_res %>%
  filter(version %in% c("v2", "v5", "v6", "v7"))

@@

by_version <- state_res %>%
  select(-date) %>%
  distinct() %>% 
  mutate(in_interval = case_when(
    exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
    exp_cases_lb  > infections  ~ "Below Interval",
    exp_cases_ub < infections ~ "Above Interval"
  )) %>%
  filter(!is.na(in_interval)) %>%
  group_by(in_interval, vlabel) %>%
  summarize(n=n()) %>%
  group_by(vlabel) %>%
  mutate(total = sum(n)) %>%
  mutate(prop=n/total)

@@

by_version %>%
  mutate(in_interval = factor(in_interval, 
                             levels = c("Above Interval",
                                        "In Interval",
                                        "Below Interval"
                                        ))) %>%
   ggplot(aes(x= fct_reorder(vlabel,prop,max), 
             fill = in_interval, y =prop)) + 
  geom_bar(stat="identity",position="stack") +
  theme_c() +
  coord_flip()+ scale_fill_manual(values=c("In Interval" = "#79D2AF",
                             "Above Interval" = "#D2AF79",
                      "Below Interval"="#7997D2"),
                       breaks = c("Below Interval",
                               "In Interval",
                               "Above Interval")) +
  labs(x="", 
       y= "Proportion",
       fill = "Covidestim Median:",
       title = "Proportion of Simulation Intervals Where Covidestim Median\nWas Within, Above, or Below the Median",
       subtitle = "Including All States") +
  theme_c() +
  theme_c(legend.position="right",
          legend.title = element_text(face="bold", size = 15),
          legend.text= element_text( size = 15),
          legend.spacing.y = unit(3, 'pt'),
          axis.text.y = element_text(size = 17, hjust=1)) +
  theme(plot.title = element_text(size = 20),
          plot.subtitle = element_text(size=18, 
                                       face='italic', 
                                       margin=margin(4,0,4,0)),
        axis.text.x=element_text(size=12),
        axis.title = element_text(size=14)) +
  scale_x_discrete(labels = (TeX(labels)))

@@

@@

names(pal) <- c(unique(subset$vlabel), "Covidestim")

@@

@@

```{r, fig.height=11, fig.height=9}

subset %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill = vlabel),
             alpha = .8,
             show.legend=FALSE) +
  geom_ribbon(aes(x = date,
                  fill = 'Covidestim',
                  ymin = infections.lo,
                  ymax = infections.hi),
              alpha = .6) +
  facet_wrap(~state, ncol=5, scales = "free_y",
              labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(axis.text.x=element_text(size=6),
          axis.text.y = element_text(size=8),
          legend.position = "top") +
  scale_fill_manual(values =pal,
                    labels = TeX(names(pal)),
                    name='',
                    breaks="Covidestim")  +
  labs(y = "Estimated Infections",
       title = paste0("Estimated Infections by State")) +
  guides(color = guide_legend(override.aes = list(linewidth =2),
                              nrow=2,
                              byrow=TRUE))

@@

@@

plt1 <- subset %>% 
  # left_join(codes) %>%
  filter(biweek == 7) %>%
  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
  theme(axis.text.y=element_text(size=10),
        axis.text.x = element_text(size=11),
        axis.title = element_text(size = 17),
        plot.subtitle = element_text(size=14),
        plot.title = element_text(size=17, face="plain")) +
  labs(y="", x="Ratio of Estimated Infections\nto Observed",
       title ="During Two-week Period\nin the Alpha Wave",
       subtitle = "March 26, 2021 through April 8, 2021")

@@

plt2 <- subset %>% 
  # left_join(codes) %>%
  filter(biweek == 13) %>%
  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
 theme(axis.text.y=element_text(size=10),
        axis.text.x = element_text(size=11),
        axis.title = element_text(size = 17),
        plot.subtitle = element_text(size=14),
        plot.title = element_text(size=17, face="plain")) +
  labs(y="", x="Ratio of Estimated Infections\nto Observed",
       title ="During Two-week Period\nin the Delta Wave",
       subtitle = "June 18, 2021 through July 1, 2021")

@@

plt3 <- subset %>% 
  # left_join(codes) %>%
  filter(biweek == 27) %>%
  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
  theme(axis.text.y=element_text(size=10),
        axis.text.x = element_text(size=11),
        axis.title = element_text(size = 17),
        plot.subtitle = element_text(size=14),
        plot.title = element_text(size=17, face="plain")) +
  labs(y="", x="Ratio of Estimated Infections\nto Observed",
       title ="During Two-week Period\nin the Omicron Wave",
       subtitle = "December 31, 2021 through January 14, 2022")

@@

plt <- plot_grid(plt1,plt2, plt3, nrow=1, align="none")

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,cache=FALSE)

@@

filter_versions <- FALSE
save_figures <- FALSE

@@

# read state-level results
targets_dir <- here("_targets")

@@

if(filter_versions){
  state_res <- state_res %>%
  filter(version %in% c("v3", "v5", "v6", "v7"))

@@

by_version <- state_res %>%
 # filter(biweek >=6) %>% 
  select(-date) %>%
  distinct() %>% 
  mutate(in_interval = case_when(
    exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
    exp_cases_lb  > infections  ~ "Below Interval",
    exp_cases_ub < infections ~ "Above Interval"
  )) %>%
  filter(!is.na(in_interval)) %>%
  group_by(in_interval, vlabel) %>%
  summarize(n=n()) %>%
  group_by(vlabel) %>%
  mutate(total = sum(n)) %>%
  mutate(prop=n/total)

@@

by_version %>%
  mutate(in_interval = factor(in_interval, 
                             levels = c("Above Interval",
                                        "In Interval",
                                        "Below Interval"
                                        ))) %>%
   ggplot(aes(x= fct_reorder(vlabel,prop,max), 
             fill = in_interval, y =prop)) + 
  geom_bar(stat="identity",position="stack") +
  theme_c() +
  coord_flip()+ scale_fill_manual(values=c("In Interval" = "#79D2AF",
                             "Above Interval" = "#D2AF79",
                      "Below Interval"="#7997D2"),
                       breaks = c("Below Interval",
                               "In Interval",
                               "Above Interval")) +
  labs(x="", 
       y= "Proportion",
       fill = "Covidestim Median:",
       title = "Proportion of Simulation Intervals Where Covidestim Median\nWas Within, Above, or Below the Median") +
  theme_c() +
  theme_c(legend.position="right",
          legend.title = element_text(face="bold", size = 15),
          legend.text= element_text( size = 15),
          legend.spacing.y = unit(3, 'pt'),
          axis.text.y = element_text(size = 17, hjust=1)) +
  theme(plot.title = element_text(size = 20),
          plot.subtitle = element_text(size=18, 
                                       face='italic', 
                                       margin=margin(4,0,4,0)),
        axis.text.x=element_text(size=12),
        axis.title = element_text(size=14)) +
  scale_x_discrete(labels = (TeX(labels)))

```{r,eval=save_figures, fig.width=14}

ggsave(here('figures/concordance-covidestim-state.pdf'))

```{r, fig.width=17}

by_version %>%
  group_by(vlabel) %>%
  filter(in_interval == "In Interval") %>%
  mutate(n_interval = n) %>%
  ggplot(aes(y=fct_reorder(vlabel,prop), x=prop)) +
  geom_text(aes(label=round(prop,3), x= prop+.03), 
             angle=0) +
  geom_bar(stat="identity", fill="#596281") +
  scale_y_discrete(breaks= unique(by_version$vlabel),
                   labels=function(x)TeX(x)) +
  scale_x_continuous(expand=c(0,0), 
                     limits=c(0,.9), n.breaks=7)  +
  theme_c(axis.text.y = element_text(size = 13, hjust=1),
          axis.title.x = element_text(size=14),
          axis.text.x=element_text(size=10)) +
  labs(y="",
       x="Proportion Containing the Covidestim Median",
       title="Proportion Containing the Covidestim Median",
       subtitle = "By Implementation")

@@

### All Versions

```{r, fig.width=18,fig.height=8}

@@

names(pal) <- c(as.character(unique(subset$vlabel)), "Covidestim")

@@

### Original Priors versus CTIS Priors (that do not vary)

state_res %>%
  filter(version %in% c("v1","v7") & state %in% sample(unique(.$state),15)) %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill=vlabel),
             alpha = .6,
             show.legend=FALSE,
             ) +
  facet_wrap(~name, 
             labeller=labeller(vlabel =as_labeller(
               TeX, default=label_parsed),
               nrow=3),
             scales="free_y") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(
          legend.position = "top",
          ) +
  viridis::scale_fill_viridis(option="magma", begin=.1,end=.9, discrete=TRUE)  +
  labs(x="",
       y = "Estimated Infections",
       title = "Simulation Intervals Over Time",
       subtitle = "By State")

@@

```{r, fig.height=14, fig.width=11}

subset %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill = vlabel),
             alpha = .8,
             show.legend=FALSE) +
  geom_ribbon(aes(x = date,
                  fill = 'Covidestim',
                  ymin = infections.lo,
                  ymax = infections.hi),
              alpha = .6) +
  facet_wrap(~name, ncol=4, scales = "free",
              labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3.5 months", 
               date_labels = "%b %Y") +
  theme_c(axis.text.x=element_text(size=5),
          axis.text.y = element_text(size=4.5),
          legend.position = "top",
          legend.text=element_text(size=12),
          strip.text.x=element_text(size=8,
                                    face='plain',
                                    color='white')) +
  scale_fill_manual(values =pal,
                    labels = TeX(names(pal)),
                    name=''
                    #,
                  # , breaks="Covidestim"
                    )  +
  labs(y = "Estimated Infections",
       title = paste0("Estimated Infections by State")) +
  guides(color = guide_legend(override.aes = list(linewidth =2),
                              nrow=2,
                              byrow=TRUE))

```{r,eval=save_figures, fig.height=14, fig.width=9}

@@

```{r,eval=save_figures, fig.height=10,fig.width=12}

@@

plt1 <- subset %>% 
  # left_join(codes) %>%
  filter(biweek == 7) %>%
  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
  theme(axis.text.y=element_text(size=10),
        axis.text.x = element_text(size=11),
        axis.title = element_text(size = 17),
        plot.subtitle = element_text(size=14),
        plot.title = element_text(size=17, face="plain")) +
  labs(y="", x="Ratio of Estimated Infections\nto Observed",
       title ="During Two-week Period\nin the Alpha Wave",
       subtitle = "March 26, 2021 through April 8, 2021") +
  xlim(0,65)

@@

plt2 <- subset %>% 
  # left_join(codes) %>%
  filter(biweek == 13) %>%
  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
 theme(axis.text.y=element_text(size=10),
        axis.text.x = element_text(size=11),
        axis.title = element_text(size = 17),
        plot.subtitle = element_text(size=14),
        plot.title = element_text(size=17, face="plain")) +
  labs(y="", x="Ratio of Estimated Infections\nto Observed",
       title ="During Two-week Period\nin the Delta Wave",
       subtitle = "June 18, 2021 through July 1, 2021") +
  xlim(0,65)

@@

plt3 <- subset %>% 
  # left_join(codes) %>%
  filter(biweek == 27) %>%
  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  theme_c() + 
  theme(axis.text.y=element_text(size=10),
        axis.text.x = element_text(size=11),
        axis.title = element_text(size = 17),
        plot.subtitle = element_text(size=14),
        plot.title = element_text(size=17, face="plain")) +
  labs(y="", x="Ratio of Estimated Infections\nto Observed",
       title ="During Two-week Period\nin the Omicron Wave",
       subtitle = "December 31, 2021 through January 14, 2022") +
  xlim(0,65)

@@

plt <- plot_grid(plt1,plt2, plt3, nrow=1, align="none", labels="auto", label_size=19)

@@

```{r,eval=save_figures, fig.height=10, fig.width=17}

ggsave(here('figures/ratio-estimated-to-observed-multiple-waves.pdf'))

subset %>% 
  # left_join(codes) %>%
  filter(biweek %in% c(7, 13, 27))  %>%
  select(-date) %>%
  distinct() %>%
  group_by(state) %>%
  mutate(ord = empirical_testpos[which(biweek==7)]) %>%
  mutate(biweek=factor(biweek)) %>%
  ggplot(aes(y=fct_reorder(state,ord),
             x=empirical_testpos,fill=biweek)) +
  geom_bar(stat="identity",position="dodge")+
  theme_c() +
  theme_c() +
  labs(title = "Testing Positivity by State",
       y="",
       x="Biweekly Testing Positivity",
       fill="") +
  scale_x_continuous(n.breaks=10)

subset %>% 
  # left_join(codes) %>%
  filter(biweek %in% c(7, 13, 27))  %>%
  group_by(biweek) %>%
  mutate(daterange= paste(format(min(date), "%B %d %Y"),
                          "to", 
                          format(max(date), "%B %d %Y")),
         daterange=factor(
           daterange,
           levels=c(
             "March 26 2021 to April 08 2021",
             "June 18 2021 to July 01 2021",
             "December 31 2021 to January 14 2022"))) %>%
  select(-date) %>%
  distinct() %>%
  group_by(state) %>%
  mutate(testrate=total/population,
         ord = testrate[which(biweek==7)]) %>%
  mutate(biweek=factor(biweek)) %>%
  ggplot(aes(y=fct_reorder(name,ord,.desc=TRUE),
             x=testrate,fill=daterange)) +
  geom_bar(stat="identity",position="dodge") +
  theme_c() +
  labs(title = "Testing Rate by State",
       y="",
       x="Biweekly Testing Rate",
       fill="") +
  scale_x_continuous(n.breaks=10)

compare_testrates <- subset %>%
  mutate(testrate=total/population) %>%
  select(-date) %>%
  distinct() %>%
  filter(biweek %in% c(7, 13, 27))  %>%
  select(biweek,state,testrate) %>%
  pivot_wider(names_from=biweek, values_from=testrate, names_prefix="biweek_") %>%
  mutate(omicron_over_alpha = biweek_27/biweek_7,
         omicron_over_delta = biweek_27/biweek_13) %>%
  select(state,contains("omicron")) %>%
  pivot_longer(c(omicron_over_alpha,omicron_over_delta)) %>%
  group_by(name) %>%
  summarize(mean_val = mean(value))

cat("The testing rate during this two-week interval during the omicron wave was", compare_testrates %>% filter(name=="omicron_over_alpha") %>% pull(mean_val), "times the testing rate during this time interval in the alpha wave.")

cat("The testing rate during this two-week interval during the omicron wave was", compare_testrates %>% filter(name=="omicron_over_delta") %>% pull(mean_val), "times the testing rate during this time interval in the delta wave.")

# Rank Ratio of Estimated to Observed Over Time

```{r,fig.height=8}

ranks <- subset %>%
  mutate(ratio = exp_cases_median/positive,
         tested = total/population) %>%
  # one observation per biweek per state
  select(biweek, date, ratio, name,tested) %>%
  group_by(biweek,ratio,name,tested) %>%
  summarize(date=min(date)) %>%
  # rank for each time interval
  group_by(biweek) %>%
  mutate(rank = rank(ratio),
         rank_tested=rank(-tested))

r <- ranks %>%
  mutate(rank_group = case_when(
    rank <= 10 ~  "Top 10",
    rank > 41 ~ "Bottom 10",
    TRUE ~ "Middle" )) %>%
  group_by(name, rank_group) %>%
  mutate(n=n()) %>%
  group_by(name) %>%
  mutate(total =n(),
         max_group = rank_group[which.max(n)]) %>%
  filter( max(n) >24) %>%
  filter(max_group != "Middle")  %>%
  ungroup()

top_10 <- r %>% filter(rank_group=="Top 10") %>%
  arrange(desc(n)) %>% pull(name) %>%
  unique() %>%paste(collapse=", ")

cat("States that consistently have the lowest ratios:", top_10)

bottom_10 <- r %>% filter(rank_group=="Bottom 10") %>%
  arrange(desc(n)) %>% pull(name) %>%
  unique() %>%paste(collapse=", ")

cat("States that consistently have the highest ratios:", bottom_10)

end_labs <- r %>%
  ungroup() %>%
  filter(date==max(date)) %>%
  mutate(date = date + days(10)) %>%
  select(name, rank, date) %>%
  ungroup()

##############################
# RANK PLOT OVER TIME
##############################
# 
# set.seed(999)
# 
# pal1 <-viridis_pal(option="viridis", end = .9, begin=.3)(3)
# pal2 <- viridis_pal(option="rocket", end = .9, begin=.3)(2)
# pal3 <- viridis_pal(option="magma", end = .9, begin=.3)(2)
# pal4 <- viridis_pal(option="mako", end = .8, begin=.3)(2)
# pal5 <- viridis_pal(option="cividis", end = .9, begin=.1)(3)
# 
# 
# rankpal <- c(pal1, pal2, pal3,pal4,pal5)
# indexes <- sample.int(length(rankpal), replace=FALSE)
# rankpal <- rankpal[indexes]

set.seed(123)

rankpal <- c("#b4ddd4", "#7b2c31", "#65d04b",
             "#da239b", "#b7d165", "#453fbc", 
             "#658114", "#af3fe8", "#ffb947",
             "#0a60a8", "#f87945", "#16d0ae")

rankpal <- c("#851657", "#be3acd", "#19a71f",
             "#824598", "#ed820a", "#BB8E37",
             "#974810", "#1f84ec", "#d02023",
             "#059dc5", "#f23387", "#16d0ae")

indexes <- sample.int(length(rankpal), replace=FALSE)
rankpal <- rankpal[indexes]

r %>%
 # filter(rank_group!="Middle") %>%
  ggplot() +
  geom_point(aes(x=date,y=rank, 
             group=name,
             color= name),
             size=.5) +
  geom_line(aes(x=date,y=rank, 
             group=name,
             color= name)) +
 # facet_wrap(~max_group) +
  ggrepel::geom_text_repel( aes(x= date-days(4),
                y = rank,
                color=name,
                label = name),
           end_labs,
           min.segment.length=2,
           nudge_y=0,
           hjust=0,
           size=3,
           direction="y",
           force_pull=9,
           box.padding=.15) +
  theme_c(legend.position="none",
          axis.text.x = element_text(angle =0,size=14)) +
  # theme(plot.subtitle =element_text(size=18),
  #       axis.title.x=element_text(size=16),
  #       axis.title.y=element_text(size=16)) +
 # scale_color_brewer(palette="Dark2")  +
  scale_color_manual(values=rankpal) +
  scale_x_date(date_breaks="2 months", date_labels = "%b %Y",
               limits = c(ymd("2021-01-01"),
                          ymd("2022-04-15"))) +
  labs(y = "Rank of Ratio",
       title = "Rank of Estimated Infections to Observed Infections Over Time",
       subtitle = "For States Ranked in the Top or Bottom 10\nfor More Than 80% of Time Intervals",
       x="Date") +
  scale_y_reverse(breaks=seq(1,51, by=9))
